---
output: github_document
---

# flowcluster

<!-- badges: start -->

[![R-CMD-check](https://github.com/Hussein-Mahfouz/flowcluster/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Hussein-Mahfouz/flowcluster/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

The goal of flowcluster is to provide minimal functionality for clustering OD desire lines (or flows). This includes: 1. Creating distance matrices between OD pairs 2. Passing distance matrices to a clustering algorithm

## Install from GitHub

You can install the development version of flowcluster from GitHub with:

``` r
#| eval=FALSE
# install.packages("devtools")
devtools::install_github("Hussein-Mahfouz/flowcluster")

```

Or, for local development, clone the repo, open the project in RStudio, and run:

```{r}
#| results: hide
devtools::load_all()
```


# Example

```{r}
library(flowcluster)
flows_sf <- flows_leeds
head(flows_sf)
plot(flows_sf["count"])
flows_clustered <- flowcluster::flowcluster(
  flows_sf
)
```

# Style

We use the tidyverse style.
To ensure your changes abide by this, run the following command before pushing changes:

```{r}
#| eval: false
styler::style_pkg()
```